---
title: "Canopy conductance"
author: "Yair Rinat""
date: "15 בנובמבר 2016"
output: html_document
---

## The canopy conductance function, including Jarvis model parameters and LAI

Kc (canopy conductance)= Kleaf (Jarvis model. function of cleaf_max,Rnrad,Rvpd,RT and RTeta) * LAI
The R in Jarvis parameters is for "relative". the values can be taken from the plots in "Physical Hydrology: Third Edition" by S. Lawrence Dingman
(p. 282)
Default values for the relative parameters are set to Rnrad=0.7, Rvpd=0.6, RT=0.8, Rteta=0.7
You can use this model to test the influence of Cleaf_max and LAI parameters on the canopy conductance (m/s).
The Kc values can be used in Penman-Montieth model.

```{r}
# The canopy_conductance function:
canopy_conductance=
  function(Cleaf_max, Rnrad=0.7, Rvpd=0.6, RT=0.8, Rteta=0.7, LAI) {

    #claculating Kc
    Kc = (Cleaf_max*Rnrad*Rvpd*RT*Rteta*LAI)
    
    #return Kc
    return(Kc)
  }
```


## Example: LAI influence analysis (while Cleaf_max=100)
```{r}
#typical LAI values:
LAI_i=c(1,5,10,20,30,40,50);
#Calculating Kc for differnt LAI values:
Kc_res=matrix(,nrow = length(LAI_i), ncol = 1);
for (i in 1:length(LAI_i)){
    Kc_res[i,1]<-(canopy_conductance(Cleaf_max = 100, LAI=LAI_i[i]));
  }

```

### plot
```{r echo=FALSE}
plot(LAI_i,Kc_res,main="canopy conductance VS LAI", xlab="LAI (m^2/m^2)", ylab="Kc (m/s)")

```

__As it can be easily seen from the model function and the plot, there is a linear relationship between a Kc and LAI. Thus, as more leaves in a given area, the higher canopy conductance you get _(!)_.__